---
title: "Calibration"
author: "Diaz Given Sangeve"
date: "2026-05-09"
output: html_document
---

# Purpose

The purpose of this R code is to calculate the steady-state values for US GDP and public debt over the sample period from Q1 1982 to Q3 2025. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Clean environment
rm(list = ls())
gc()

# Loading Packages 
if(!require(pacman)) { install.packages("pacman"); require(pacman)}
pacman::p_load(fredr, lubridate, tidyverse)
```

# 1. API configuration

To configure the data pipeline, you must first obtain a free API key by registering for an account on the St. Louis Fed (FRED) website and requesting a key through your account dashboard. To use this key securely without hardcoding your credentials into the script, save it in your local environment file (typically an .Renviron file) as FRED_API_KEY="your_key_here". When the script executes, Sys.getenv("FRED_API_KEY") automatically retrieves this value from your system, keeping your credentials secure and making the code easily reproducible across different machines.

```{r API}
# Set your FRED API key
api_key <- Sys.getenv("FRED_API_KEY")
```

# 2. Data Retrieval

```{r data-fetch}
start_qtr <- as.Date("1982-01-01")
end_qtr <- as.Date("2025-07-01")

# Fetch National Accounts and Debt 
debt_gdp <- fredr(series_id = "GFDEGDQ188S", observation_start = start_qtr, observation_end = end_qtr)
real_gdp <- fredr(series_id = "GDPC1", observation_start = start_qtr, observation_end = end_qtr)

# Nominal Government Expenditures and Nominal GDP
gov_exp <- fredr(series_id = "W019RCQ027SBEA", observation_start = start_qtr, observation_end = end_qtr)
nom_gdp <- fredr(series_id = "GDP", observation_start = start_qtr, observation_end = end_qtr)

# Rates, Prices, and Population (forcing quarterly aggregation)
fed_funds <- fredr(series_id = "FEDFUNDS", observation_start = start_qtr, observation_end = end_qtr, frequency = "q", aggregation_method = "avg")
cpi <- fredr(series_id = "CPIAUCSL", observation_start = start_qtr, observation_end = end_qtr, frequency = "q", aggregation_method = "avg")
pop <- fredr(series_id = "LFWA64TTUSM647S", observation_start = start_qtr, observation_end = end_qtr, frequency = "q", aggregation_method = "avg")
```

# 3. Per-Capita Transformations and Calibration

```{r calibration}
# Ensure alignment (assuming data frames are identically ordered by date)
# 1. Transform to Real Per Capita series
real_gdp_pc <- real_gdp$value / pop$value

# (Note: Gov Expenditure and Debt ratios are scale-free, so we can use them directly or construct their real per capita equivalents if needed for standard deviations later)
g_y_series <- gov_exp$value / nom_gdp$value

# 2. Compute the Steady-State Ratios
b_y <- mean(debt_gdp$value, na.rm = TRUE) / 100
g_y <- mean(g_y_series, na.rm = TRUE)


# 6. Compile Results
calibrated_results <- tibble(
  Parameter = c("Debt-to-GDP (b_y)",
                "Gov Expenditure-to-GDP (g_y)"
                ),
  Value = c(b_y, g_y)
)

print(calibrated_results)
```

