---
output:
  md_document:
    variant: markdown_github
---

# Purpose

Providing the r code required to obtain the statistics for our Macroeconomics project.


```{r}

rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
library(tidyverse)
if(!require(readxl)) { install.packages("readxl"); require(readxl)} 
library(readxl)
if(!require(mFilter)) { install.packages("mFilter"); require(mFilter)} 
library(mFilter)
if(!require(knitr)) { install.packages("knitr"); require(knitr)} 
library(knitr)

list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))
```

# Loading Data

```{r}
Data <- read_excel('/Users/diaz/Documents/Masters/Coursework/First Semester/Macroeconomics/Project/Macro Project/data/Macro_Data.xlsx',
                   sheet = "Logged Data")  # change to your sheet name

head(Data)
colnames(Data)
```

# Applying HP filter to logged quantity variables

We do this because the model's business cycle statistics are computed from cyclical component, which are the deviations from trend. We thus need to apply the HP filter to each log-transformed series (which we can't do in excel directly)

```{r}
hp_gdp  <- hpfilter(Data$Log_GDP,  freq = 1600)
hp_cons <- hpfilter(Data$Log_Cons, freq = 1600)
hp_debt <- hpfilter(Data$Log_Debt, freq = 1600)
hp_inf  <- hpfilter(Data$Inflation, freq = 1600)
hp_ffr  <- hpfilter(Data$FFR,       freq = 1600)
```

# Extracting cyclical components of the data

```{r}
cycle_gdp  <- hp_gdp$cycle
cycle_cons <- hp_cons$cycle
cycle_debt <- hp_debt$cycle
cycle_inf  <- hp_inf$cycle
cycle_ffr  <- hp_ffr$cycle
```

# Obtaining the Standard Deviations

```{r}
sd_gdp  <- sd(cycle_gdp)  * 100
sd_cons <- sd(cycle_cons) * 100
sd_debt <- sd(cycle_debt) * 100
sd_inf  <- sd(cycle_inf)  * 100
sd_ffr  <- sd(cycle_ffr)  * 100
```

# Obtaining the Correlations

```{r}
corr_cons <- cor(cycle_cons, cycle_gdp)
corr_debt <- cor(cycle_debt, cycle_gdp)
corr_inf  <- cor(cycle_inf,  cycle_gdp)
corr_ffr  <- cor(cycle_ffr,  cycle_gdp)
```

# Build the results table

```{r}
results_table <- data.frame(
  Variable = c("Output", "Household Consumption", "Inflation (CPI)", 
               "Public Debt", "Federal Funds Rate"),
  
  Std_Dev = c(round(sd_gdp,  3),
              round(sd_cons, 3),
              round(sd_inf,  3),
              round(sd_debt, 3),
              round(sd_ffr,  3)),
  
  Corr_with_Output = c(1.000,
                       round(corr_cons, 3),
                       round(corr_inf,  3),
                       round(corr_debt, 3),
                       round(corr_ffr,  3))
)
```

# Print results as a clean table

```{r}
colnames(results_table) <- c("Variable", "Standard Deviation (%)", 
                              "Correlation with Output")
kable(results_table, align = c("l", "c", "c"))
```

